Something about Performance, 2D Component of NEMO. (TBC)

### 2016-01-07 nemolite2d_perf.tex
nemolite2d paper: update reference to top500 list in intro

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HE" in caps to complete the first word.
\IEEEPARstart{T}{he} challenge presented to the developers of
scientific software by the drive towards Exascale computing is
considerable. With power consumption becoming the overriding design
constraint, CPU clock speeds are falling and the complex,
multi-purpose compute core is being replaced by multiple, simpler
cores. This philosophy can be seen at work in the rise of so-called
accelerator based machines in the Top 500 List~\cite{top500} of
supercomputers: five of the top-ten machines in the November 2014 list
make use of Intel Xeon Phi's or NVIDIA GPUs. Four of the remaining
five machines in the top ten are IBM BlueGene/Qs, the CPU of which has
hardware support for running 64 threads.

@@

\subsection{The `nemolite2d' Program}
For this work we have used a  program

{\bf Description of the structure of the program at the Algorithm
  level so that it links in to Profile data in results section.}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HE" in caps to complete the first word.
\IEEEPARstart{T}{he} challenge presented to the developers of
scientific software by the drive towards Exascale computing is
considerable. With power consumption becoming the overriding design
constraint, CPU clock speeds are falling and the complex,
multi-purpose compute core is being replaced by multiple, simpler
cores. This philosophy can be seen at work in the rise of so-called
accelerator-based machines in the Top 500 List~\cite{top500} of
supercomputers: four of the top-ten machines in the November 2015 list
make use of Intel Xeon Phi's or NVIDIA GPUs. Two of the remaining six
machines are IBM BlueGene/Qs, the CPU of which has hardware support
for running 64 threads.

@@

\subsection{The `nemolite2d' Program}

For this work we have used a program "nemolite2d" developed by
ourselves.  “nemolite2d” is a vertically-averaged version of NEMO
(Nucleus for European Modelling of the Ocean), which only keeps its
dynamical part. The whole model system is represented through one
continuity equation (for the update of the sea-surface height) and two
vertically-integrated momentum equations (for the two velocity
components, respectively).

(add above three equations in .tex file).

The external forcing includes surface wind stress, bottom friction,
and open boundary barotropic forcing. A lateral slip boundary
condition is applied along the coast lines. The open boundary
conditions can be set as a clipped or a Flather's radiation
condition~\cite{flather76}. The bottom friction takes a semi-implicit
form for the sake of model stability. As done in the original version
of NEMO, a constant or Smagorinsky horizontal viscosity coefficient is
used for the horizontal viscosity term.

The traditional Arakawa C structured grid is employed here for the
discretisation of the computational domain. A two-dimensional integer
array is used to identify the different parts of the computational
domain; it has the value of one for ocean, zero for land and minus
one for ocean cells outside of the computational domain. This array
serves as a mark to identify ocean cells, land cells, solid boundaries
and open boundaries.

For the sake of simplicity, the explicit Eulerian forward time
stepping method is implemented here, except that the bottom friction
takes a semi-implicit form.  The Coriolis force can be set in explicit
or implicit form. The advection term is computed with a first-order
upwind scheme.

The order of the model computation is as follows:
\begin{enumerate}
\item Set the initial conditions (water depth, sea surface height, velocity)
\item integrate the continuity equation for the new sea surface height
\item update the different terms in the right hand side of the momentum equations (advection, Coriolis forcing(if set in explicit form), pressure gradient, and, horizontal viscosity
\item update the velocity vectors by summing up the values in 3), and implicitly with the bottom friction and Coriolis forcing (if set in implicit form)
\item apply the boundary conditions on to the open boundary and solid boundary cells.
\end{enumerate}
